\PYGEOMETRY{}: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-09-14 public/paper/pyg4ometry.tex
Paper update

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material radiation transport (RT) programmes these 
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}. 
All these codes are based on the Monte Carlo technique but each code either has a particular specialism, simulation methodology or target user community.  
RT Monte Carlo (RTMC) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical imaging and therapy and space radiation environment  simulation are some examples. A fundamental requirement of all of the codes is to supply a computer-readable description of the 
physical three-dimensional geometry that  the particles are passing through.  The creation of geometry files is typically a very time-consuming 
activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no standard geometry format used across MCRT 
codes, with each MCRT code typically using its own unique format. The requirement of a geometry system that allows the conversion between files prepared for different 
codes will enable cross-checks of the physics processes in different particle transport codes. Typically an MCRT code user will not have geometry in FLUKA 
and Geant4 for example. The file formats used for geometry are generally focused on the computational efficiency of a particle tracking task and not ease of use of a user. In addition to the creation of geometry files for RT programs, usually computer-aided design (CAD) files exist for systems which 
need to be simulated. The fundamental geometric representations in CAD files are usually not amenable for MCRT programs.  For these reasons it is 
advantageous to create a software tool that allows particle transport code users to rapidly develop error-free geometry files, convert between common MCRT 
program geometry formats and load CAD models.

@@

\begin{figure*}[hbt!]
  \normalsize
  \centering
  \includegraphics[width=0.7\textwidth]{./diagrams/workflow.pdf}
  \caption{\label{fig:workflow}Schematic of \PYGEOMETRY{} workflow, showing (a) the different input file formats, (b) Python processing, (c) output file targets and (d) RTMC codes which use the geometry. }
\end{figure*}

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis program, can load, display and manipulate GDML-like geometry. Tools exist to convert CAD files to GDML, open source examples are GUIMesh \cite{GUIMesh_Pinto}, commercial products include ESABASE2 \cite{ESABASE2} and FASTRAD \cite{FASTRAD}. There are also tools from the fusion and 
neutronics community that can convert CAD geometry into formats usable by RTMC codes examples include DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}.  
In principle CAD software can export shape data to STL (or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. In practice, using a lot of CAD models is difficult if a CAD model is comprised of a large number of parts. Exporting, assigning material to and placing the STL components into the RTMC can be a cumbersome task. The existent set of software does not provide a complete set of tools to efficiently create complex geometries.

@@

%%%%%%%%%%%%%%%%% Boundary representation (BREP)
Boundary representation consists of two parts, topology and geometry. Topological elements are faces, edges and vertices and the corresponding
geometrical elements are surfaces, curves and points. No current RTMC applications use native file formats employed by CAD systems. The conversion of 
CAD BREP formats for loading in RTMC applications is typically performed via a tessellated format, although it is possible to decompose BREP descriptions
 to bounded or infinite mathematical surfaces and subsequently solids as used in CSG descriptions. This type of conversion is complex and error-prone, 
 although recent progress has been made \colorbox{yellow}{CITEXXX}.

%%%%%%%%%%%%%%%%% Tessellated polygons
Solid volumes can be defined using triangular, quadrilateral or tetrahedral meshes. Numerous formats exist to describe meshes, the ubiquitous being STL and 
more modern variants being PLY, OBJ to name a few. For solids with curved faces a tessellated mesh will always give an approximate description. As the mesh 
deviation distance from the solid decreases the number of polygons increases and with it the memory consumption and execution time of the RTMC simulation.

@@

%%%%%%%%%%%%%%%%%  Internal data representation
The internal data representation follows closely the structure of GDML. A \verb|Registry| class aggregates Python ordered dictionaries that are  used to store the main 
elements of a GDML file. As a \PYGEOMETRY{} user specifies geometry the registry is correspondingly updated. When a user is finished with the geometry, the registry is written to disk as a GDML file.
\verb|Registry| instances can be aggregated to form a composite geometry or volumes can be removed and added.

@@

%%%%%%%%%%%%%%%%% Volume bounding
If a user is creating and placing multiple daughter volumes within a {\em mother} volume then it is the users responsibility to create a solid which fully 
encompasses the daughter volumes. Overlaps between daughter volumes and the mother can be detected to, but it is desirable \colorbox{yellow}{It's not just desirable, it's absolutely necessary isn't it?  And I'm not sure what the word ``efficiently'' means precisely in this context.} to have a mother volume 
shape that efficiently holds all of its daughters.

@@

%%%%%%%%%%%%%%%%% Adding properties to volumes
Geometry needs to be augmented with other information for a complete RTMC simulation. Often other attributes need to be 
assigned to regions of space for example material definition, magnetic field or optical properties. These physical properties 
can be used to define the visualisation attributes of a volume.

\subsection{Overlap detection}
All RTMC codes cannot handle spatial overlap between two geometric objects and will have ill-defined behaviour when tracking particles  
in such a situation.  A key feature of \PYGEOMETRY{} is the detection of potential overlaps in a way which is most useful to the user, it does this by performing an intersection operation between solid instances and determining if the resultant mesh is empty. Figure~\ref{fig:overlap} shows the potential overlaps, (a) protrusion of a daughter from the mother, (b)  finite volume intersection between two daughters  and (c) an overlap where two daughters share a face. If the resulting intersection 
is non-null then the overlaps can be displayed side-by-side in the visualisation. 
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=8cm]{./diagrams/overlap.pdf}
\caption{Schematic of the three different types of overlap between daughters of a mother logical volume.}
\label{fig:overlap}
\end{center}
\end{figure}

@@

For a logical volume with $n_{\rm daughters}$ physical volumes, assuming meshes for solids have an average number of faces $n$, clearly this algorithm 
has complexity $\sim \mathcal{O}(n^2 n_{\rm daughters}^2)$. This might seem like a prohibitive computational cost, but worth it considering the potential 
waste if small overlaps are present in the final RTMC simulation and large amounts of cluster CPU time is wasted. This algorithm clearly favours geometry descriptions which have a high degree of logical volume reuse. Due to the discrete nature of triangular meshes it is not possible to have perfect detection 
of overlaps, especially when curved surfaces are considered. The deviation of the meshes created from a solid can be controlled by the user reducing the 
chances of missing a potential overlap, but the algorithm presented here cannot capture all overlap cases. The overlap detection algorithm can 
present the potential overlaps quickly and easily to the user, thus significantly aiding their modelling process.

%%%%%%%%%%%%%%%%% Graphical user interface
% A basic Qt based graphical user interface is also available with \PYGEOMETRY{}. If a model has a large number of volumes defined in a hierarchical tree structure, it is % cumbersome for the user to find the logical or physical volume.

@@

\section{GDML to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 
solids can be mapped to a FLUKA \verb|REGION|. Some solids in Geant4 directly map to 
a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between Geant4 solids and the bodies used to compose a FLUKA region.
A region is a region of space which is formed of bodies and the operators + (intersection), $-$ 
(complement) and | (union), also parentheses to indicate the order of operations. FLUKA 
has 20 bodies, listed in Table~\ref{tab:Fluka2Geant4}  and, in general, infinite extent bodies have 
benefits over finite ones. Key for conversion are XY-, XZ-, YZ-Planes (XYP, XZP, YZP), arbitrary 
plane (PLA), Z-axis aligned cylinder (ZCC), Z-axis aligned elliptical cylinder (ZEC), 
sphere (SPH), truncated right-angle cone (TRC) and general quadric surface (QUA). 
\begin{figure}
\begin{center}
\includegraphics[width=0.9\columnwidth]{./model-scene/CuboidalChamber.pdf}
\caption{Example conversion of a simple GDML geometry to FLUKA, left panel:  
the original model in \PYGEOMETRY{}, right panel: the converted FLUKA geometry 
viewed in FLAIR. The example is a sector bend dipole electromagnet.}
\label{fig:gdml-flair}
\end{center}
\end{figure}

@@

There are some important issues which must be considered to perform an accurate conversion. 
A Geant4 logical volume solid might need to be created from a FLUKA region composing of many zones, 
so $R_1 = +z_1 | +z_2$, similarly a void is required to locate daughter volume placements which again 
might also be converted to a region of multiple zones $R_2= +z_3 | +z_4$. So creating the FLUKA region is then 
\begin{eqnarray}
R_1 - R_2 	& = & (+z_1 | +z_2) - ( +z_3 | +z_4) 			\\
			& = & (+z_1 - z_3 - z_4) \;  | \; (+z_2 - z_3 - z_4)	
\label{eqn:setDiff}
\end{eqnarray}
Geant4 has the Boolean solids associated with difference, union and intersection, so in addition to 
Equation~\ref{eqn:setDiff}, both $R_1 \cup R_2$ and $R_1 \cap R_2$ is required in FLUKA notation, so 
\begin{eqnarray}
R_1 \cup R_2 	& = & (+z_1 | +z_2)  \cup ( +z_3 | +z_4) \\
			& = & +z_1 | +z_2 |  +z_3 | +z_4
\label{eqn:setUnion}
\end{eqnarray}
and 
\begin{eqnarray}
R_1 \cap R_2 	& = & (+z_1 | +z_2) \cap ( +z_3 | +z_4) \\
			& = & +z_1 +z_3  \; | +z_1 +z_4 \; | +z_2 +z_3 \; | +z_2 +z_4
\label{eqn:setIntersection}
\end{eqnarray}

@@

\subsection{Discussion}
Overall the conversion to FLUKA input format from GDML is quite 
advanced and stable. Relatively large experimental simulations 
have been converted from GDML to FLUKA and have been producing 
simulation results. The conversion described still requires a user to
understand how geometry is specified in both Geant4 and FLUKA. For 
example the subtraction of many non-primitive body daughter volumes from 
a mother will create a very complex non-DNF region which is inefficient when viewed in 
FLAIR and used for simulation in FLUKA. A geometry designer can avoid this by 
restricting daughter volumes solids in Geant4 that can be expressed in DNF simply,
for example a cuboid.  If \PYGEOMETRY{} is used during the creation of
 geometry then multiple codes can be targeted without additional user effort. 
 \colorbox{yellow}{There are still a few outstanding technical  ssues, which should be discussed. }

@@

@@

\begin{enumerate}
%\item Degenerate body look up 
%\item Bodies with/without transformation 
% \item Non-convex solids, convex decomposition (TwistedBox, TwistedTrap, TwistedTrd, TwistedTube, Tessellated)
% \item Scaled solid, Placement scale 
% \item Replica, Division, Parameterised
\item Recursive application of \ref{eqn:setDiff}, \ref{eqn:setUnion}, \ref{eqn:setIntersection}, \ref{eqn:logicalSubtraction} speed up using meshes
% \item Issues with tessellated solids 
\end{enumerate}

@@

STEP and IGES files can be loaded into \PYGEOMETRY{}, via an interface 
based on FreeCAD. FreeCAD is an open source CAD/CAM program, which 
in turn is based on OpenCASCADE. FreeCAD allows for scripting in Python 
and acts as a simple to use interface to OpenCASCADE.  A STEP CAD file 
could be considered as a hierarchical tree of part features and part assemblies, 
where a part assembly is a collection of part features. A part feature can be used 
to create a triangular mesh which can be used to create a \PYGEOMETRY{} tessellated solid. The placement 
of the part feature is extracted from the STEP file and used to create an 
appropriate physical volume. Assignment of materials and visualisation 
attributes must be performed by  the user after conversion to GDML as rarely 
will CAD/CAM packages included the detailed information required for RTMC 
codes. Listing~\ref{lst:pythonCADLoading} shows 
how \PYGEOMETRY{} can load step files.
\begin{lstlisting}[caption={A simple \PYGEOMETRY{} Python script to load a STEP file.},label={lst:pythonCADLoading}, language=Python]
    reader = pyg4ometry.freecad.Reader("CadFileName.step")
    reader.relabelModel()
    reader.convertFlat()
    logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model 
part names can contain characters which are not allowed in Python dictionaries so need to be replaced by \pyinline{relabelModel}. 
CAD models might also have a hierarchy of parts and assemblies, these are converted without this structure by \pyinline{convertFlat}.

@@

\section{Quality assurance}
The source and manual code for \PYGEOMETRY{} is stored in a git repository (\url{https://bitbucket.org/jairhul/pyg4ometry}), 
where a public issue tracker for users is hosted to report problems or bugs with the code. The manual is created using SPHINX a 
Python documentation generator. \PYGEOMETRY{} uses mature packages available for Python as dependences. 
\PYGEOMETRY{} has  two submodules that require compilation \pyinline{pyg4ometry.pycsg} and \pyinline{pyg4ometry.pycgal} 
in C/C++. \PYGEOMETRY{}, package dependencies and extensions are installed using \pyinline{setuptools}. All aspects of 
\PYGEOMETRY{} are routinely checked using unit tests, of which 543 tests are available for the code, which have 84\% coverage.

@@

The {\bf conversion} which would most dramatically enhance the creation of geometry is CAD to Geant4/FLUKA
without use use of a triangular or tetrahedral mesh. There are methods to decompose BREP solids to Geant4 and 
FLUKA like CSG geometry \colorbox{yellow}{CITEXXX MCCAD}. The FreeCAD/OpenCASCADE interface
combined with the Geant4 and FLUKA Python API present in \PYGEOMETRY{} will allow for the creation 
of CAD B-REP decomposition algorithms.

There is a strong relationship between \PYGEOMETRY and Geant4 and to a lesser extent between 
\PYGEOMETRY{} and FLUKA. \PYGEOMETRY{} can be used to a testing ground for ideas prior to 
implementation in Geant4 or FLAIR. An example of this the VTK visualisation system implemented in 
\PYGEOMETRY{}, which could be used in Geant4 to rendering of Boolean solids which frequently 
fail in the Geant4 OpenGL viewer. This would involve using CGAL meshing in the \pyinline{G4Polyhedron}
class. There is also

>>>>

%% main text
\section{Introduction} \label{sec:introduction}
There are numerous different software codes to simulate the passage of particles through material radiation transport (RT) programmes these 
include MCNP~\cite{Mcnp_Werner}, FLUKA~\cite{Fluka_Ferrari,Fluka_Bohlen}, Geant3~\cite{Geant3_Brun} and Geant4~\cite{Geant4_Agostinelli}. 
All these codes are based on the Monte Carlo technique but each code either has a particular specialism, simulation methodology or target user community.  
 Monte Carlo RT (MCRT) simulations have diverse uses including shielding calculations for radiological protection, detector performance, medical 
imaging and therapy and space radiation environment  simulation are some examples. A fundamental requirement of all of the codes is to supply a 
computer-readable description of the  physical three-dimensional geometry that  the particles are passing through.  The creation of geometry files is 
typically a very time-consuming activity and the simulation validity and performance is directly dependent on the quality of the geometry. There is no 
standard geometry format used across MCRT codes, with each MCRT code typically using its own unique format. The requirement of a geometry system 
that allows the conversion between files prepared for different codes will enable cross-checks of the physics processes in different particle transport
 codes. Typically an MCRT code user will not have geometry in FLUKA and Geant4 for example. The file formats used for geometry are generally focused 
 on the computational efficiency of a particle tracking task and not ease of use of a user. In addition to the creation of geometry files for RT programs, usually 
 computer-aided design (CAD) files exist for systems which need to be simulated. The fundamental geometric representations in CAD files are usually not 
 amenable for MCRT programs.  For these reasons it is advantageous to create a software tool that allows particle transport code users to rapidly develop 
 error-free geometry files, convert between common MCRT program geometry formats and load CAD models.

@@

\begin{figure*}[hbt!]
  \normalsize
  \centering
  \includegraphics[width=0.7\textwidth]{./diagrams/workflow.pdf}
  \caption{\label{fig:workflow}Schematic of \PYGEOMETRY{} workflow, showing (a) the different input file formats, (b) Python processing, (c) output file targets and 
  (d) MCRT codes which use the geometry. }
\end{figure*}

There are existing codes that have functionality similar to \PYGEOMETRY{}. ROOT~\cite{Brun:1997pa}, the high-energy physics data analysis program, 
can load, display and manipulate GDML-like geometry. Tools exist to convert CAD files to GDML, open source examples are GUIMesh \cite{GUIMesh_Pinto}, commercial 
products include ESABASE2 \cite{ESABASE2} and FASTRAD \cite{FASTRAD}. There are also tools from the fusion and neutronics community that can convert CAD 
geometry into formats usable by MCRT codes examples include DAGMC~\cite{DAGMC} and McCAD~\cite{McCad}. In principle CAD software can export shape data to STL 
(or other similar mesh formats), which can be used by Geant4~\cite{poole2012acad}. In practice, using a lot of CAD models is difficult if a CAD model is comprised of a 
large number of parts. Exporting, assigning material to and placing the STL components into the MCRT can be a cumbersome task. The existent set of software does not 
provide a complete set of tools to efficiently create complex geometries.

@@

%%%%%%%%%%%%%%%%% Boundary representation (BREP)
Boundary representation consists of two parts, topology and geometry. Topological elements are faces, edges and vertices and the corresponding
geometrical elements are surfaces, curves and points. No current MCRT applications use native file formats employed by CAD systems. The conversion of 
CAD BREP formats for loading in MCRT applications is typically performed via a tessellated format, although it is possible to decompose BREP descriptions
 to bounded or infinite mathematical surfaces and subsequently solids as used in CSG descriptions. This type of conversion is complex and error-prone, 
 although recent progress has been made \colorbox{yellow}{CITEXXX}.

%%%%%%%%%%%%%%%%% Tessellated polygons
Solid volumes can be defined using triangular, quadrilateral or tetrahedral meshes. Numerous formats exist to describe meshes, the ubiquitous being STL and 
more modern variants being PLY, OBJ to name a few. For solids with curved faces a tessellated mesh will always give an approximate description. As the mesh 
deviation distance from the solid decreases the number of polygons increases and with it the memory consumption and execution time of the MCRT simulation.

@@

%%%%%%%%%%%%%%%%%  Internal data representation
The internal data representation follows closely the structure of GDML. A \verb|Registry| class aggregates Python ordered dictionaries that are  used to store the main 
elements of a GDML file. As a \PYGEOMETRY{} user specifies geometry the registry is correspondingly updated. When a user is finished with the geometry, the registry 
is written to disk as a GDML file.\verb|Registry| instances can be aggregated to form a composite geometry or volumes can be removed and added.

@@

%%%%%%%%%%%%%%%%% Volume bounding
If a user is creating and placing multiple daughter volumes within a {\em mother} volume then it is the users responsibility to create a solid which fully 
encompasses the daughter volumes. Overlaps between daughter volumes and the mother can be detected to, but it is desirable 
% \colorbox{yellow}{It's not just desirable, it's absolutely necessary isn't it?  And I'm not sure what the word ``efficiently'' means precisely in this context.} 
to have a mother volume shape that efficiently holds its daughters. This is not always the case for example assembly volumes.

@@

%%%%%%%%%%%%%%%%% Adding properties to volumes
Geometry needs to be augmented with other information for a complete MCRT simulation. Often other attributes need to be 
assigned to regions of space for example material definition, magnetic field or optical properties. These physical properties 
can be used to define the visualisation attributes of a volume.

\subsection{Overlap detection}
All MCRT codes cannot handle spatial overlap between two geometric objects and will have ill-defined behaviour when tracking particles  
in such a situation.  A key feature of \PYGEOMETRY{} is the detection of potential overlaps in a way which is most useful to the user, it does this by 
performing an intersection operation between solid instances and determining if the resultant mesh is empty. Figure~\ref{fig:overlap} shows the potential 
overlaps, (a) protrusion of a daughter from the mother, (b)  finite volume intersection between two daughters  and (c) an overlap where two daughters 
share a face. If the resulting intersection is non-null then the overlaps can be displayed side-by-side in the visualisation. 
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=8cm]{./diagrams/overlap.pdf}
\caption{Schematic of the three different types of overlap between daughters of a mother logical volume.}
\label{fig:overlap}
\end{center}
\end{figure}

@@

For a logical volume with $n_{\rm daughters}$ physical volumes, assuming meshes for solids have an average number of faces $n$, clearly this algorithm 
has complexity $\sim \mathcal{O}(n^2 n_{\rm daughters}^2)$. This might seem like a prohibitive computational cost, but worth it considering the potential 
waste if small overlaps are present in the final MCRT simulation and large amounts of cluster CPU time is wasted. This algorithm clearly favours geometry 
descriptions which have a high degree of logical volume reuse. Due to the discrete nature of triangular meshes it is not possible to have perfect detection 
of overlaps, especially when curved surfaces are considered. The deviation of the meshes created from a solid can be controlled by the user reducing the 
chances of missing a potential overlap, but the algorithm presented here cannot capture all overlap cases. The overlap detection algorithm can 
present the potential overlaps quickly and easily to the user, thus significantly aiding their modelling process.

@@

\section{GDML to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 
solids can be mapped to a FLUKA \verb|REGION|. Some solids in Geant4 directly map to 
a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between Geant4 solids and the bodies used to compose a FLUKA region.
A region is a region of space which is formed of bodies and the operators $+$ (intersection), $-$ 
(complement) and $\:|\:$ (union), also parentheses to indicate the order of operations. FLUKA 
has 20 bodies, listed in Table~\ref{tab:Fluka2Geant4}  and, in general, infinite extent bodies have 
benefits over finite ones. Key for conversion are XY-, XZ-, YZ-Planes (XYP, XZP, YZP), arbitrary 
plane (PLA), Z-axis aligned cylinder (ZCC), Z-axis aligned elliptical cylinder (ZEC), 
sphere (SPH), truncated right-angle cone (TRC) and general quadric surface (QUA). 
\begin{figure}
\begin{center}
\includegraphics[width=0.9\columnwidth]{./model-scene/CuboidalChamber.pdf}
\caption{Example conversion of a simple GDML geometry to FLUKA, left panel:  
the original model in \PYGEOMETRY{}, right panel: the converted FLUKA geometry 
viewed in FLAIR. The example is a sector bend dipole electromagnet.}
\label{fig:gdml-flair}
\end{center}
\end{figure}

@@

There are some important issues which must be considered to perform an accurate conversion. 
A Geant4 logical volume solid might need to be created from a FLUKA region composing of many zones, 
so $R_1 = +z_1\: | +z_2$, similarly a void is required to locate daughter volume placements which again 
might also be converted to a region of multiple zones $R_2= +z_3 \: | +z_4$. So creating the FLUKA region is then 
\begin{eqnarray}
R_1 - R_2 	& = & (+z_1 \: | +z_2) - ( +z_3 \: | +z_4) 			\\
			& = & (+z_1 - z_3 - z_4) \;  | \; (+z_2 - z_3 - z_4)	
\label{eqn:setDiff}
\end{eqnarray}
Geant4 has the Boolean solids associated with difference, union and intersection, so in addition to 
Equation~\ref{eqn:setDiff}, both $R_1 \cup R_2$ and $R_1 \cap R_2$ is required in FLUKA notation, so 
\begin{eqnarray}
R_1 \cup R_2 	& = & (+z_1 \: | +z_2)  \cup ( +z_3 \: | +z_4) \\
			& = & +z_1 \: | +z_2 |  +z_3 \: | +z_4
\label{eqn:setUnion}
\end{eqnarray}
and 
\begin{eqnarray}
R_1 \cap R_2 	& = & (+z_1 \: | +z_2) \cap ( +z_3 \: | +z_4) \\
			& = & +z_1 +z_3  \; | +z_1 +z_4 \; | +z_2 +z_3 \; | +z_2 +z_4
\label{eqn:setIntersection}
\end{eqnarray}

@@

\subsection{Discussion}
Overall the conversion to FLUKA input format from GDML is quite 
advanced and stable. Relatively large experimental simulations 
have been converted from GDML to FLUKA and have been producing 
simulation results. The conversion described still requires a user to
understand how geometry is specified in both Geant4 and FLUKA. For 
example the subtraction of many non-primitive body daughter volumes from 
a mother will create a very complex non-DNF region which is inefficient when viewed in 
FLAIR and used for simulation in FLUKA. A geometry designer can avoid this by 
restricting daughter volumes solids in Geant4 that can be expressed in DNF simply,
for example a cuboid.  If \PYGEOMETRY{} is used during the creation of
 geometry then multiple codes can be targeted without additional user effort. 
There are still a few outstanding technical issues with the conversion, which are discussed 
in this section.

@@

In general, recursive application of Equations \ref{eqn:setDiff},
\ref{eqn:setUnion}, \ref{eqn:setIntersection} and \ref{eqn:logicalSubtraction} 
can result in very complex regions when converting from Geant4/GDML to 
FLUKA. The complexity of the final region expression can be compounded 
if transformed to DNF. The final region Boolean expression can be simplified by 
evaluating using the meshes for the surfaces, only retaining terms which 
do not evaluation to a null mesh.

@@

@@

STEP and IGES files can be loaded into \PYGEOMETRY{}, via an interface 
based on FreeCAD. FreeCAD is an open source CAD/CAM program, which 
in turn is based on OpenCASCADE. FreeCAD allows for scripting in Python 
and acts as a simple to use interface to OpenCASCADE.  A STEP CAD file 
could be considered as a hierarchical tree of part features and part assemblies, 
where a part assembly is a collection of part features. A part feature can be used 
to create a triangular mesh which can be used to create a \PYGEOMETRY{} tessellated solid. The placement 
of the part feature is extracted from the STEP file and used to create an 
appropriate physical volume. Assignment of materials and visualisation 
attributes must be performed by  the user after conversion to GDML as rarely 
will CAD/CAM packages included the detailed information required for MCRT 
codes. Listing~\ref{lst:pythonCADLoading} shows 
how \PYGEOMETRY{} can load step files.
\begin{lstlisting}[caption={A simple \PYGEOMETRY{} Python script to load a STEP file.},label={lst:pythonCADLoading}, language=Python]
    reader = pyg4ometry.freecad.Reader("CadFileName.step")
    reader.relabelModel()
    reader.convertFlat()
    logical = reader.getRegistry().getWorldVolume()
\end{lstlisting}
Compared to other file readers, two additional steps are required \pyinline{relabelModel} and \pyinline{convertFlat}. CAD model 
part names can contain characters which are not allowed in Python dictionaries so need to be replaced by \pyinline{relabelModel}. 
CAD models might also have a hierarchy of parts and assemblies, these are converted without this structure by \pyinline{convertFlat}.

@@

\section{Quality assurance}
The source and manual code for \PYGEOMETRY{} is stored in a git repository (\url{https://bitbucket.org/jairhul/pyg4ometry}), 
where a public issue tracker for users is hosted to report problems or bugs with the code. The manual is created using Sphinx a 
Python documentation generator. \PYGEOMETRY{} uses mature packages available for Python as dependences. 
\PYGEOMETRY{} has  two submodules that require compilation \pyinline{pyg4ometry.pycsg} and \pyinline{pyg4ometry.pycgal} 
in C/C++. \PYGEOMETRY{}, package dependencies and extensions are installed using \pyinline{setuptools}. All aspects of 
\PYGEOMETRY{} are routinely checked using unit tests, of which 543 tests are available for the code, which have 84\% coverage. 
The unit tests also serve as minimal examples to help users understand the code operaton.

@@

The {\bf conversion} which would most dramatically enhance the creation of geometry is CAD to Geant4/FLUKA
without use use of a triangular or tetrahedral mesh. There are methods to decompose BREP solids to Geant4 and 
FLUKA like CSG geometry \colorbox{yellow}{CITEXXX MCCAD}. The FreeCAD/OpenCASCADE interface
combined with the Geant4 and FLUKA Python API present in \PYGEOMETRY{} will allow for the creation 
of CAD BREP decomposition algorithms.

There is a strong relationship between \PYGEOMETRY and Geant4 and to a lesser extent between 
\PYGEOMETRY{} and FLUKA. \PYGEOMETRY{} can be used to a testing ground for ideas prior to 
implementation in Geant4 or FLAIR. An example of this the VTK visualisation system implemented in 
\PYGEOMETRY{}, which could be used in Geant4 to rendering of Boolean solids which frequently 
fail in the Geant4 OpenGL viewer. This would involve using CGAL meshing in the \pyinline{G4Polyhedron}
class. %There is also
